A Computational Account of Default-Mode Function\\ by Control Theory and Reinforcement Learning

### 2016-09-29 paper/paper.tex
grammar

\section{Reinforcement-Learning as a computational model for DMN function}
The main idea is to show that the DMN has biological substrates (gradient
computation, backprop, experience replay, etc.) for all the
major components of an efficient RL algorithm, and all these components
integrate in a plausible way.
\subsection{General Reinforcment-Learning (RL) theory}
In a general reinforcement-learning (RL) framework, an \textit{agent}
interacts with an unknown environment in trial-and-error fashion by taking
\textit{actions / decisions} $a_t$ which trigger changes in the
\textit{state} of the environment
$s_t \rightarrow s_{t+1}$, accompanied with a real-valued \textit{reward}
$r_t$ (or regret!, since it can be negative). This reward should be thought
of as an instantaneous satisfaction accompanying the execution of
an action. Here the subscript $t$ denotes time, and can be discrete or
continuous.
%% The unknown environment $\mathcal E$ is generally taken as stochastic, but having some kind of
%% memory.
A popular abstract model for the scenario are so-called \textit{Markov Decision Processes (MDPs)}.
This is simply a quintuplet $(\mathcal S, \mathcal A, r, \rho, \mu)$ where
\begin{itemize}
\item $\mathcal S$ is the set of states, e.g $\mathcal S = \{\text{happy}, \text{sad}\}$.
\item $\mathcal A$ is the set of actions, e.g $\mathcal A = \{\text{read}, \text{run},
  \text{laugh}, \text{empathize}, \text{empathize}\}.$
\item $r : \mathcal S \times \mathcal A \rightarrow \mathbb R$, so that $r(s, a)$ is the instant reward for taking action $a$ at state $s$.
\item $\mu: \mathcal S \rightarrow [0, 1]$ is the prior probability on the states so that
  $\mu(s)$ is the probability that the environment starts-off in state $s$.
  \item $\rho : \mathcal S \times \mathcal A \times \mathcal S \rightarrow [0, 1]$, so that
    $\rho(s,a,s') = \mathbb P(s'|s, a)$ is the probability of moving to state $s'$ if action $a$ is taken from state $s$. In addition, one requires that such transitions be Markovian.
\end{itemize}

@@

The behavior of the agent is governed by some kind of \textit{policy}, which maps states to
sampling probabilities over candidate actions to perform from that state. The  goal of RL is then
adapt this policy so as to  maximize the cumulative rewards.

@@

By construction, in general, maximizing the expectation in  \eqref{eq:qval} is intractable due to
the enormous number of possible trajectories to consider.
One popular simplification that removes this issue is to only consider the sub-family of \textit{deterministic policies}, which map each state to the best single best action which should be taken at that state, and therefore define functions from states to actions.

@@

\subsection{Q-learning with value approximation and replay}
As already mentioned in the previous section, Q-learning optimizes over the class of
deterministic policies of the form \eqref{eq:qlearning}.
At state $s \in \mathcal S$, the agent takes an action $a \in \mathcal A$
% (for example, $a$ = ``Read'', or ``Run'', or ``Laugh'', or ``Sympathize'', or ``Empathize'', etc.)
by sampling from its current policy matrix, and collects a reward $r$,
% (measured on their internal currency of worldly values ?),
and the environment transitions to a new $s' \in \mathcal S$. At the end of
this cycle, a new \textit{experience} $e = (s,a,r,s')$ is produced; this represents an exemplar behavior of the agent and is recorded in replay memory:
$\mathcal D \leftarrow \mathcal D \cup \{e\}$.
Now, replay consists in sampling (uniform or importance-weighted ?)  mini-batches of experiences
$(s, a, r, s') \sim \mathcal U(D)$ from the replay memory $\mathcal D$ and trying to
approximate
the would-be $Q$-value at $(s, a)$ as predicted by Bellman's equation \eqref{eq:bellman}, namely
$r + \gamma \max_{a'} Q(s', a'|\theta^Q)$, with the output of a parametrized regression model $(s,a,r,s') \mapsto {Q}(s, a|\theta^Q)$, where $Q(.|\theta^Q)$ is an approximator for the $Q$-value operator,  parameterized by $\theta^Q$. Computing an optimal policy then corresponds to finding the parameters $\theta^Q$ which minimize the following loss function
\begin{equation}
  \mathcal L(\theta^Q)
  % = \mathbb E_{(s, a, r, s') \sim U(\mathcal D)}[(Q(s, a|\theta^Q) - Q(s',a'|\theta^Q))^2]
  = \mathbb E_{(s, a, r, s') \sim U(\mathcal D)}[(Q(s, a|\theta^Q) - y)^2],
  \label{eq:oracle}
  \end{equation}
where
$ y = r + \gamma \max_{a' \in \mathcal A} Q(s', a'|\theta^Q) = r + \gamma Q(s', \pi(s')|\theta^Q)$ is the prediction target.
% While $y_t$ is also dependent on $\theta^Q$, this is typically ignored.

>>>>

\section{Reinforcement-Learning as a computational model for DMN function}
The main idea is to show that the DMN has biological substrates (gradient
computation, backprop, experience replay, etc.) for all the
major components of an efficient RL algorithm, and all these components
integrate in a plausible way.
\subsection{General Reinforcment-Learning (RL) theory}
In a general reinforcement-learning (RL) framework, an \textit{agent}
interacts with an unknown environment in trial-and-error fashion by taking
\textit{actions / decisions} $a_t$ which trigger changes in the
\textit{state} of the environment
$s_t \rightarrow s_{t+1}$, accompanied with a real-valued \textit{reward}
$r_t$ (or regret!, since it can be negative). This reward should be thought
of as an instantaneous satisfaction accompanying the execution of
an action. Here the subscript $t$ denotes time, and can be discrete or
continuous.
%% The unknown environment $\mathcal E$ is generally taken as stochastic, but having some kind of
%% memory.
A popular abstract model for the scenario are so-called \textit{Markov Decision Processes (MDPs)}.
This is simply a quintuplet $(\mathcal S, \mathcal A, r, \rho, \mu)$ where
\begin{itemize}
\item $\mathcal S$ is the set of states, e.g $\mathcal S = \{\text{happy}, \text{sad}\}$.
\item $\mathcal A$ is the set of actions, e.g $\mathcal A = \{\text{read}, \text{run},
  \text{laugh}, \text{empathize}, \text{empathize}\}.$
\item $r : \mathcal S \times \mathcal A \rightarrow \mathbb R$, so that $r(s, a)$ is the instant reward for taking action $a$ in state $s$.
\item $\mu: \mathcal S \rightarrow [0, 1]$ is the prior probability on the states so that
  $\mu(s)$ is the probability that the environment starts-off in state $s$.
  \item $\rho : \mathcal S \times \mathcal A \times \mathcal S \rightarrow [0, 1]$, so that
    $\rho(s,a,s') = \mathbb P(s'|s, a)$ is the probability of moving to state $s'$ if action $a$ is taken from state $s$. In addition, one requires that such transitions be Markovian.
\end{itemize}

@@

The behavior of the agent is governed by some kind of \textit{policy}, which maps states to
sampling probabilities over candidate actions to perform from thin state. The  goal of RL is then
adapt this policy so as to  maximize the cumulative rewards.

@@

By construction, in general, maximizing the expectation in  \eqref{eq:qval} is intractable due to
the enormous number of possible trajectories to consider.
One popular simplification that removes this issue is to only consider the sub-family of \textit{deterministic policies}, which map each state to the best single best action which should be taken at thin state, and therefore define functions from states to actions.

@@

\subsection{Q-learning with value approximation and replay}
As already mentioned in the previous section, Q-learning optimizes over the class of
deterministic policies of the form \eqref{eq:qlearning}.
in state $s \in \mathcal S$, the agent takes an action $a \in \mathcal A$
% (for example, $a$ = ``Read'', or ``Run'', or ``Laugh'', or ``Sympathize'', or ``Empathize'', etc.)
by sampling from its current policy matrix, and collects a reward $r$,
% (measured on their internal currency of worldly values ?),
and the environment transitions to a new $s' \in \mathcal S$. At the end of
this cycle, a new \textit{experience} $e = (s,a,r,s')$ is produced; this represents an exemplar behavior of the agent and is recorded in replay memory:
$\mathcal D \leftarrow \mathcal D \cup \{e\}$.
Now, replay consists in sampling (uniform or importance-weighted ?)  mini-batches of experiences
$(s, a, r, s') \sim \mathcal U(D)$ from the replay memory $\mathcal D$ and trying to
approximate
the would-be $Q$-value at $(s, a)$ as predicted by Bellman's equation \eqref{eq:bellman}, namely
$r + \gamma \max_{a'} Q(s', a'|\theta^Q)$, with the output of a parametrized regression model $(s,a,r,s') \mapsto {Q}(s, a|\theta^Q)$, where $Q(.|\theta^Q)$ is an approximator for the $Q$-value operator,  parameterized by $\theta^Q$. Computing an optimal policy then corresponds to finding the parameters $\theta^Q$ which minimize the following loss function
\begin{equation}
  \mathcal L(\theta^Q)
  % = \mathbb E_{(s, a, r, s') \sim U(\mathcal D)}[(Q(s, a|\theta^Q) - Q(s',a'|\theta^Q))^2]
  = \mathbb E_{(s, a, r, s') \sim U(\mathcal D)}[(Q(s, a|\theta^Q) - y)^2],
  \label{eq:oracle}
  \end{equation}
where
$ y = r + \gamma \max_{a' \in \mathcal A} Q(s', a'|\theta^Q) = r + \gamma Q(s', \pi(s')|\theta^Q)$ is the prediction target.
% While $y_t$ is also dependent on $\theta^Q$, this is typically ignored.
